Hierarchical Region-Network Sparsity for\\ High-Dimensional Inference in Brain Imaging

### 2016-02-05 paper/icml2016/example_paper.tex
remove padding

\section{Introduction}
% sparsity
Many quantitative scientific domains underwent a
recent passage from the classical ``long data" regime to
the high-dimensional ``wide data" regime.
% the high-dimensional regime (i.e., ``wide data") \cite{jordan2015massive}.
Also in the brain imaging domain,
many contemporary methods for acquiring brain signals yield
more variables per observation than
total observations per data sample.
This $n \ll p$ scenario challenges various statistical methods from
classical statistics.
For instance,
estimating generalized linear models without additional assumptions
yields an underdetermined system of equations.
%
Many such ill-posed estimation problems
have benefited from
\textit{sparsity} assumptions
\cite{hastie2015statistical}.
They act as regularizer by encouraging zero coefficients
in model selection.
Sparse supervised and unsupervised
learning algorithms have proven to yield
statistical relationships that can be readily
estimated, reproduced, and interpreted
\cite{giraud2014introduction}.
%
Moreover, \textit{structured sparsity} can impose
domain knowledge on the 
statistical estimation,
thus shrinking and selecting variables guided by
expected data distributions
\cite{bach2012optimization}.
Such restrictions to statistical complexity
are an attractive plan of attack
for the $\textgreater$100,000 variables per brain map.
Yet, what generally accepted neurobiological structure lends itself
to exploitation using structured sparsity priors?

% specialization & integration
Concepts on human brain organization have long been torn
between the two extremes
\textit{functional specialization} and \textit{functional integration}.
Functional specialization emphasizes that microscopically distinguishable
brain regions are responsible for distinct classes of computational 
processes
\cite{kanwisher2010functional}.
Conversely, functional integration emphasizes that brain computation
is enabled by a complex interplay between these
distinct brain regions \cite{sporns14nn}.
%
Local
infrastructure
and global connectivity profiles are thought to go hand-in-hand
to realize neural processes.
%
However,
probably no existing brain analysis method acknowledges that
both functional design principles are inextricably involved
in the realization of mental operations.

Functional specialization has long been
explored and interpreted based on different research methods.
%
For instance,
single-cell recordings and microscopic tissue examination
revealed the
segregation of the occipital visual cortex into V1, V2, V3, V3A/B, and V4
regions
\cite{hubel1962receptive, zeki1978functional}.
Tissue lesion of the mid-fusiform gyrus of the visual system,
in turn,
was frequently reported to impair
recognition of others' identity from faces
\cite{iaria2008contrib}.
%
As a crucial common point,
all these methodological approaches
yield neuroscientific findings
that are naturally interpreted according to
non-overlapping, discrete region compartments
as the basic architecture of brain organization.
% INTEGRAL
It is more recent
that the main interpretational focus has shifted
from circumscribed regions to network stratifications
in systems neuroscience \cite{yuste2015}.
%
Besides analyses of
electrophysiological oscillations
% \cite{buzsaki2004neuronal}
and
graph-theoretical properties,
% graph-theoretical properties \cite{bullmore2009complex},
studies of
functional connectivity correlation \cite{buckner2013opportunities} and
independent component analysis (ICA) \cite{beckmann2005}
became the workhorses of network discovery
in neuroimaging.
%
As a common point of this second set of methods,
interpretation of findings naturally embraces
cross-regional integration by
overlapping network compartments
as the basic architecture of brain organization,
in stark constrast to methods examining regional specialization.

@@

One can read off the formula that decreasing \(\alpha\) leads to less 
penalization of brain networks and thus to tendentially fully active groups
and dense brain maps. If at the same time \(\beta\) is increased to 
induce group sparsity, then only the structure of brain regions encoded
by \(\mathcal G\) is present. Conversely, if \(\beta\) is chosen 
sufficiently small, and \(\alpha\) increased,
then the structure will come from \(\mathcal H\), leading to the
selection of brain networks rather than regions.

>>>>

\section{Introduction}
% sparsity
Many quantitative scientific domains underwent a
recent passage from the classical regime (i.e., ``long data")  to
the high-dimensional regime (i.e., ``wide data").
% the high-dimensional regime (i.e., ``wide data") \cite{jordan2015massive}.
Also in the brain imaging domain,
many contemporary methods for acquiring brain signals yield
more variables per observation than
total observations per data sample.
This $n \ll p$ scenario challenges various statistical methods from
classical statistics.
For instance,
estimating generalized linear models without additional assumptions
yields an underdetermined system of equations.
%
Many such ill-posed estimation problems
have benefited from
\textit{sparsity} assumptions
\cite{hastie2015statistical}.
They act as a 
regularizer and can be used for model selection.
Sparse supervised and unsupervised
learning algorithms have proven to yield
statistical relationships that can be readily
estimated, reproduced, and interpreted
\cite{giraud2014introduction}.
%
Generally, \textit{structured sparsity} can impose
domain knowledge on the 
statistical estimation,
thus shrinking and selecting variables guided by
expected data distributions
\cite{bach2012optimization}.
Such restrictions to statistical complexity
are an attractive plan of attack
for the $\textgreater$100,000 variables per brain map.
Yet, what generally accepted neurobiological structure lends itself
to exploitation using structured sparsity priors?

% specialization & integration
Concepts on human brain organization have long been torn
between the two extremes
\textit{functional specialization} and \textit{functional integration}.
Functional specialization emphasizes that microscopically distinguishable
brain regions are responsible for distinct classes of computational 
processes
\cite{kanwisher2010functional}.
Conversely, functional integration emphasizes that brain function
is enabled by a complex interplay between these
distinct brain regions \cite{sporns14nn}.
%
Local
infrastructure
and unique global connectivity profiles are thought to go hand-in-hand
to realize brain function.
%
However,
probably no existing brain analysis method acknowledges that
both functional design principles are inextricably involved
in the realization of mental operations.

Functional specialization has long been
explored and interpreted based on different research methods.
%
For instance,
single-cell recordings and microscopic examination
revealed the
segregation of the occipital visual cortex into V1, V2, V3, V3A/B, and V4
regions
\cite{hubel1962receptive, zeki1978functional}.
Tissue lesion of the mid-fusiform gyrus of the visual system,
in turn,
was frequently reported to impair
recognition of others' identity from faces
\cite{iaria2008contrib}.
%
As a crucial common point,
all these methodological approaches
yield neuroscientific findings
that are naturally interpreted according to
non-overlapping, discrete region compartments
as the basic architecture of brain organization.
% INTEGRAL
It is more recent
that the main interpretational focus has shifted
from circumscribed regions to network stratifications
in systems neuroscience \cite{yuste2015}.
%
Besides analyses of
electrophysiological oscillations
% \cite{buzsaki2004neuronal}
and
graph-theoretical properties,
% graph-theoretical properties \cite{bullmore2009complex},
studies of
functional connectivity correlation \cite{buckner2013opportunities} and
independent component analysis (ICA) \cite{beckmann2005}
became the workhorses of network discovery
in neuroimaging.
%
As a common point of this second set of methods,
interpretation of findings naturally embraces
cross-regional integration by
overlapping network compartments
as the basic architecture of brain organization,
in stark constrast to methods examining regional specialization.

@@

Decreasing \(\alpha\) leads to less 
penalization of brain networks and thus to tendentially fully active groups
and dense brain maps. If at the same time \(\beta\) is increased to 
induce group sparsity, then only the structure of brain regions encoded
by \(\mathcal G\) is present. Conversely, if \(\beta\) is chosen 
sufficiently small, and \(\alpha\) increased,
then the structure will come from \(\mathcal H\), leading to the
selection of brain networks rather than regions.
